### 2014-10-09 Studies/2013_PG_County/Report/PG_Co_Fire.tex
PG Co: update report and figures

@@

\section{Geometry}
\label{geom}
The structure involved in this fire incident was a one story, single-family structure with a basement. Due to the surrounding topography (the ground slopes from the front to rear of the structure), the basement on the rear side of the structure was above grade. The structure was wood-framed platform-framed with a gable roof. \footnote{Platform-framed construction means that floor joists are completely covered with a sub-floor. This forms a platform for which walls are built from.} The structure had a first floor footprint of 9.14~m (30~ft) by 7.77~m (25.5~ft) with the basement being slightly smaller, 9.14~m (30~ft) by 7.16~m (23.5~ft), as show in Fig.~\ref{fig:simp_geom}. This results in a first floor area of 71~m$^2$ (764~ft$^2$) and a basement floor area of 65~m$^2$ (700~ft$^2$). The interior walls and ceiling were constructed from nominal lumber covered by 12.7~mm (0.5~in) gypsum board~\cite{PGCounty2013}. There was a single flight of stairs in the center of the structure which connected the first floor to the basement. Access to the structure could be made on the front side of the first floor and the rear side of the basement.

\begin{figure}[!ht]
\centering
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}r}
\includegraphics[width=.7\textwidth]{../Figures/Basement} \\
\includegraphics[width=.7\textwidth]{../Figures/First_Floor}
\end{tabular*}
\caption[Plan view of basement (top) and first (bottom) floor of structure.]{Plan view of basement (top) and first (bottom) floor of structure. Stairs and key windows and doors are identified using measurements collected by NIST.}
\label{fig:simp_geom}
\end{figure}

\begin{figure}[!ht]
\centering
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}r}
\includegraphics[width=5in]{../Figures/PGCo_Front} \\ 
\includegraphics[width=5in]{../Figures/PGCo_Rear}
\end{tabular*}
\caption{Photographs showing the front (top) and rear (bottom) exterior of the structure after the incident.}
\label{fig:exterior}
\end{figure}

@@

\section{Ventilation}
\label{Vents}
The simulations in this study account for changes in the ventilation due to a combination of fire department operations (opening doors, breaking windows), fire acting on the structure (fire breaching windows/doors) and wind (restricting flow, closing doors). Three ventilation scenarios were considered in this study. Each scenario followed the timeline of events outlined in Table~\ref{tab:vents}. The difference in each scenario is the wind boundary condition. Simulations were run at 0~m/s (0~mph), 4.47~m/s (10~mph), and 8.94~m/s (20~mph) wind velocities with a flow direction from the rear toward the front of the structure. Based on local weather conditions, mean wind velocities ranged between 4.47~m/s (10~mph) and 8.94~m/s (20~mph). The case without wind was run to control for the impact of wind on the flow of hot gases within the structure.

@@

@@

\section{Numerical Mesh}
\label{mesh}
For the simulation, a measure of how well the flow field is resolved can be estimated by using the non-dimensional expression $D^*/\dx$. Here, $D^*$ is the characteristic fire diameter, $\dx$ is the nominal size of a mesh cell, and $\dQ$ is the total heat release rate of the fire:
\begin{equation}
D^* = \left(
     \frac{\dQ}{\rho_\infty \, c_p \, T_\infty \, \sqrt{g}}
     \right)^\frac{2}{5} 
\label{eq:mesh}
\end{equation}   
From the FDS User's Guide~\cite{FDS_Users_Guide}, the characteristic fire diameter is related to the characteristic fire size via the
relation $Q^* = (D^*/D)^{5/2}$. Here, $D$ is the physical diameter of the base of the fire specified in the simulation. Following Eq.~\ref{eq:mesh} and using a grid cell size of 10~cm, the characteristic fire diameter to cell size ($D^*/\dx$) ratio is 12 for the 1.5~MW bathroom fire and 20 for the 6.75~MW basement fire. Based on validation work performed for the U.S. Nuclear Regulatory Commission, $D^*/\dx$ values ranged between 4 and 16~\cite{NUREG_1824} and produced results that were adequate for engineering calculations. The grid resolution used in this model is within or exceeds typical engineering $D^*/\dx$ values.

As discussed in Section~\ref{geom}, the structure has dimensions of 9.14~m (30~ft) by 7.62~m (25~ft) and a total height of 7.31~m (24~ft). To sufficiently resolve the wind boundary condtion, the computational domain was extended beyond the volume of the structure. The large domain led to the use of two separate grid resolutions, a 10~cm resolution and a 20~cm resolution. The 10~cm resolution encompasses the structure with a total volume 14~m (46~ft) by 12~m (40ft) by 8~m (26~ft) that required 1.8 million computational cells. As a result, all of the input model geometry lengths, ventilation openings (holes, doors, and windows), and fire areas will snap to the nearest 10~cm.

Figure~\ref{fig:geom_grid} shows the front of the structure rendered in Smokeview with the 10~cm mesh. The domain was divided into 16 equally sized meshes (each containing 150,000 grid cells) that could be processed in parallel, which reduced the amount of required calculation time to approximately 1.5~days. Figure~\ref{fig:mult_mesh} shows the entire structure within the computational domain that has been divided into multiple meshes. Table~\ref{tab:mod_param} shows a summary of the model input parameters for the simulations that were conducted as part of this study.

>>>>

\section{Geometry}
\label{geom}
The structure involved in this fire incident was a one story, single-family structure with a basement. The structure was platform-framed wood construction with a gable roof. \footnote{Platform-framed construction means that floor joists are completely covered with a sub-floor. This forms a platform for which walls are built from.} The structure had a first floor footprint of 9.14~m (30~ft) by 7.77~m (25.5~ft) with the basement being slightly smaller, 9.14~m (30~ft) by 7.16~m (23.5~ft), as show in Fig.~\ref{fig:simp_geom}. This resulted in a first floor area of 71~m$^2$ (764~ft$^2$) and a basement floor area of 65~m$^2$ (700~ft$^2$). The interior walls were composed of gypsum board over insulated wood framing.~\cite{PGCounty2013}. There was a single flight of stairs in the center of the structure which connected the first floor to the basement.

The structure was built on a sloped landscape such that on the front side of the structure the first floor was located above street level but the basement was below street level (Fig.~\ref{fig:slope}). On the rear side of the structure, the slope allowed the basement to be a walkout. Therefore, access to the structure could be made on the front side of the first floor and the rear side of the basement.

\begin{figure}[!ht]
\includegraphics[width=.7\textwidth]{../Figures/PGCo_Slope} \\
\caption[Photograph of terrain.]{Photograph taken from front side street level showing the slope down to basement on the rear side of the structure.}
\label{fig:slope}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=.7\textwidth]{../Figures/Basement} \\
\vspace{0.2in}
\includegraphics[width=.7\textwidth]{../Figures/First_Floor}
\caption[Plan view of basement (top) and first (bottom) floor of structure.]{Plan view of basement (top) and first (bottom) floor of structure. Stairs and key windows and doors are identified using measurements collected by NIST.}
\label{fig:simp_geom}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/PGCo_Front} \\ 
\vspace{0.1in}
\includegraphics[width=5in]{../Figures/PGCo_Rear}
\caption{Photographs showing the front (top) and rear (bottom) exterior of the structure after the incident.}
\label{fig:exterior}
\end{figure}

On-scene photographs and post-incident reports were used to estimate the type of fuel, fire size, and ventilation openings on the structure.

@@

@@

\section{Ventilation}
\label{Vents}
The simulations in this study account for changes in the ventilation due to a combination of fire department operations (opening doors, breaking windows), fire acting on the structure (fire breaching windows/doors) and wind (restricting flow, closing doors). The ventilation times and ventilation areas represent our best understanding of the incident. The times of ventilation changes for doors and windows are provided in Table~\ref{tab:vents}.

During the time at which the firefighters exited the structure or were rescued from the structure, the state of some windows and doors in the structure was known. The state of the vents (opened versus closed) resulted in the establishment of a flow path in the interior stairwell leading from the basement to the first floor. However, the exact time of operation for other doors/windows was not known. Therefore, the following doors were assumed to be open at the start of the simulation: 1) the basement door, and 2) the basement windows on the left rear corner of the structure. The states of the front door and front window were specified at certain times based on observations (i.e. forcible entry, wind effects, forced window breakage) from post-incident reports.

@@

In addition to the state of openings within the structure, three wind scenarios were considered in this study. Each scenario followed the timeline of events outlined in Table~\ref{tab:vents}. The difference in each scenario is the wind boundary condition. Simulations were run with 0~m/s (0~mph), 4.5~m/s (10~mph), and 8.9~m/s (20~mph) wind velocities, all with a flow direction from the rear toward the front of the structure. Based on local weather conditions, mean wind velocities ranged between 4.5~m/s (10~mph) and 8.9~m/s (20~mph) with gusts up to 13.4~m/s (30~mph)~\cite{PGCounty2013}. The case without wind was run to control for the impact of wind on the flow of hot gases within the structure.

@@

\section{Numerical Mesh}
\label{mesh}
For the simulation, a measure of how well the flow field is resolved can be estimated by using the non-dimensional expression $D^*/\dx$. Here, $D^*$ is the characteristic fire diameter, $\dx$ is the nominal size of a mesh cell, and $\dQ$ is the total heat release rate of the fire:
\begin{equation}
D^* = \left(
     \frac{\dQ}{\rho_\infty \, c_p \, T_\infty \, \sqrt{g}}
     \right)^\frac{2}{5} 
\label{eq:mesh}
\end{equation}   
From the FDS User's Guide~\cite{FDS_Users_Guide}, the characteristic fire diameter is related to the characteristic fire size via the
relation $Q^* = (D^*/D)^{5/2}$. Here, $D$ is the physical diameter of the base of the fire specified in the simulation. Following Eq.~\ref{eq:mesh} and using a grid cell size of 10~cm (3.9~in), the characteristic fire diameter to cell size ($D^*/\dx$) ratio is 12 for the 1.5~MW bathroom fire and 20 for the 6.75~MW basement fire. Based on validation work performed for the U.S. Nuclear Regulatory Commission, $D^*/\dx$ values ranged between 4 and 16~\cite{NUREG_1824} and produced results that were adequate for engineering calculations. The grid resolution used in this model is within or exceeds typical engineering $D^*/\dx$ values.

As discussed in Section~\ref{geom}, the structure has dimensions of 9.14~m (30~ft) by 7.62~m (25~ft) and a total height of 7.31~m (24~ft). To sufficiently resolve the wind boundary condtion, the computational domain was extended beyond the volume of the structure. The large domain led to the use of two separate grid resolutions, a 10~cm (3.9~in) resolution and a 20~cm (7.8~in) resolution. The 10~cm resolution encompasses the structure with a total volume 14~m (46~ft) by 12~m (40ft) by 8~m (26~ft) that required a total of 1.8 million computational cells. As a result, all of the input model geometry lengths, ventilation openings (holes, doors, and windows), and fire areas will snap to the nearest 10~cm.

Figure~\ref{fig:geom_grid} shows the front of the structure and surrounding domain rendered in Smokeview with the 10~cm  and 20~cm meshes. The domain was divided into 16 equally sized meshes (each containing 150,000 grid cells) that could be processed in parallel, which reduced the amount of required calculation time to approximately 1.5~days. Figure~\ref{fig:mult_mesh} shows the entire structure within the computational domain that has been divided into multiple meshes. Table~\ref{tab:mod_param} shows a summary of the model input parameters for the simulations that were conducted as part of this study.
